---
title: "Assignment 3"
author: "Sara Østergaard"
date: "30/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
pacman::p_load(
  tidyverse,
  here,
  posterior,
  cmdstanr,
  brms, 
  boot, 
  loo,
  patchwork
)

```


## Load data
Social conformity data from schizophrenia patients
```{r}
sc <- read_csv('sc_schizophrenia.csv')

# there are NAs in the rating so we remove those rows
sc <- na.omit(sc)

sc <- sc %>% mutate(
  # ID above 200 is comparison group (0), ID below 200 is schizophrenia group (1)
  'Group' = ifelse(ID > 200, 0, 1),
  # change in rating between round 1 and round 2
  'Change' = SecondRating - FirstRating,
  # feedback difference between own rating and other rating
  'Feedback' = OtherRating - FirstRating
)

# subset of the data
sc <- sc %>% filter(ID %in% c(201, 202, 203))
```

**The data contains the following columns:**
- FaceID: an identifier of the specific face rated
- ID: an identifier of the participant
- Trial_Round1: in which trial the face was presented (during the first exposure)
- Trial_Round2: in which trial the face was presented (during the second exposure)
- FirstRating: the trustworthiness rating (1-8) given by the participant BEFORE seeing other ratings   
- OtherRating: the trustworthiness rating (1-8) given by others
- SecondRating: the trustworthiness rating (1-8) given after seeing the others (at second exposure)
- RT_Round1: time taken to produce the first rating of trustworthiness
- RT_Round2: time taken to produce the second rating of trustworthiness
*columns we have added*
- Group: 0 is comparison group, 1 is schizophrenia group
- Change: the difference between the second and the first rating
- Feedback: the difference between other rating and own first rating

## Explore and visualize data

```{r}
sc %>% group_by(ID) %>% 
  summarize('n' = n())
# all participants have 153 trials

sc %>% group_by(Group) %>% 
  summarize('n' = n()/153,
            'mean_firstRating' = mean(FirstRating),
            'sd_firstRating' = sd(FirstRating),
            'mean_secondRating' = mean(SecondRating),
            'sd_secondRating' = sd(SecondRating),
            'mean_change' = mean(Change),
            'sd_change' = sd(Change),
            'mean_feedback' = mean(Feedback),
            'sd_feedback' = mean(Feedback))
# there are a 4 more participants in the control group compared to patient group
# ratings looks fairly similar accross groups


```

```{r}
ggplot(sc, aes(Feedback, Change)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm) +
  facet_wrap("Group") + 
  theme_bw()

sc %>% filter(Group == 1) %>% ggplot(aes(Feedback, Change)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm) +
  facet_wrap("ID") + 
  theme_bw()

sc %>% filter(Group == 0) %>% ggplot(aes(Feedback, Change)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = lm) +
  facet_wrap("ID") + 
  theme_bw()
```


```{r}
face_mean <- sc %>% group_by(FaceID, Group) %>% 
  summarize(
    'mean_firstRating' = mean(FirstRating),
    'sd_firstRating' = sd(FirstRating),
    'mean_secondRating' = mean(SecondRating),
    'sd_secondRating' = sd(SecondRating)
  )

ggplot(data = sc) + 
  geom_bar(aes(FaceID, FirstRating), 
           position = "dodge", stat = "summary", fun.y = "mean") + 
  theme_bw()

ggplot(data = sc) + 
  geom_bar(aes(FaceID, SecondRating), 
           position = "dodge", stat = "summary", fun.y = "mean") + 
  theme_bw()
```

```{r}
ggplot(sc, aes(FirstRating, Feedback)) + 
  geom_point(alpha = 0.1) + theme_bw() + facet_wrap('Group')

ggplot(sc, aes(SecondRating, Feedback)) + 
  geom_point(alpha = 0.1) + theme_bw() + facet_wrap('Group')
```

## Simple Bayes
First we will fit a simple Bayes model to the data, where FirstRating is source 1 and OtherRating is source 2

```{r}
# load model
file <- file.path('simpleBayes.stan')

mod <- cmdstan_model(file, cpp_options = list(stan_threads=TRUE))
```

```{r}
# data frames with data for the stan model
df_choice <- sc %>% subset(select=c(ID, SecondRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = SecondRating)

df_s1 <- sc %>% subset(select=c(ID, FirstRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = FirstRating)

df_s2 <- sc %>% subset(select=c(ID, OtherRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = OtherRating)
```


```{r}
# data for the stan model
n_trials <- max(sc$Trial_Round1)
n_agents <- length(unique(sc$ID))

data <- list(
  trials = n_trials,
  agents = n_agents,
  choice = as.matrix(df_choice[, 2:(n_participants+1)]),
  Source1 = as.matrix(df_s1[, 2:(n_participants+1)])/10,
  Source2 = as.matrix(df_s2[, 2:(n_participants+1)])/10
)

# call stan
samples_simple <- mod$sample(
  data = data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 1000,
  max_treedepth = 20,
  adapt_delta = 0.99
)
```

```{r}
samples_simple

draws_df <- as_draws_df(samples_simple$draws())
```

```{r}
ggplot(draws_df, aes(.iteration, sigma, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()
```


## Weighted Bayes
Now we will fit a weighted Bayes model to the data, where FirstRating is source 1 and OtherRating is source 2

```{r}
# load model
file <- file.path('weightedBayes.stan')

mod <- cmdstan_model(file, cpp_options = list(stan_threads=TRUE))
```

```{r}
# data frames with data for the stan model
df_choice <- sc %>% subset(select=c(ID, SecondRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = SecondRating)

df_s1 <- sc %>% subset(select=c(ID, FirstRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = FirstRating)

df_s2 <- sc %>% subset(select=c(ID, OtherRating, Trial_Round1)) %>% 
  pivot_wider(names_from = ID, values_from = OtherRating)
```


```{r}
# data for the stan model
n_trials <- max(sc$Trial_Round1)
n_participants <- length(unique(sc$ID))

data <- list(
  trials = n_trials,
  agents = n_participants,
  choice = as.matrix(df_choice[, 2:(n_participants+1)]),
  Source1 = as.matrix(df_s1[, 2:(n_participants+1)])/10,
  Source2 = as.matrix(df_s2[, 2:(n_participants+1)])/10
)

# call stan
samples_weighted <- mod$sample(
  data = data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 1000,
  max_treedepth = 20,
  adapt_delta = 0.99
)
```

```{r}
samples_weighted$summary()

draws_df <- as_draws_df(samples_weighted$draws())
```

```{r}
ggplot(draws_df, aes(.iteration, w1M, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()

ggplot(draws_df, aes(.iteration, w2M, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()

ggplot(draws_df, aes(.iteration, sigma, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()

```


```{r}
weight1_update <- ggplot(draws_df) +
  geom_density(aes(w1M), fill='blue', alpha=0.5) + 
  geom_density(aes(w1M_prior), fill='red', alpha=0.5) + 
  theme_bw()

weight2_update <- ggplot(draws_df) +
  geom_density(aes(w2M), fill='blue', alpha=0.5) + 
  geom_density(aes(w2M_prior), fill='red', alpha=0.5) + 
  theme_bw()

weight1_update + weight2_update
```



## model comparison
We will compare the two models: simple bayes and weighted bayes

```{r}
loo_simple <-samples_simple$loo(cores=3)
loo_weighted <- samples_weighted$loo(cores=3)

plot(loo_simple)
plot(loo_weighted)

```

```{r}
loo_compare(loo_simple, loo_weighted)
loo_model_weights(list(loo_simple, loo_weighted))
```




