---
title: "assignment n"
author: "Sara Østergaard"
date: "16/3/2022"
output: html_document
---

During this assignment you will have to pick two (or more) models (e.g. Random Bias, Imperfect Memory, WSLS) and analyze one the Matching Pennies datasets available for the course (see end of instructions)

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
pacman::p_load(
  tidyverse,
  here,
  posterior,
  cmdstanr,
  brms, 
  boot, 
  loo
)
```

## load data
Data with cogsci students
```{r}
df <- read_csv('mp_students_22.csv')%>% 
  filter(BotStrategyN == -1 | BotStrategyN == -2) %>%   # we will only look at WSLS and random bias bots
  mutate(BotStrategy = ifelse(BotStrategyN == -1, 'RandomBias', 'WSLS'))

df <- df %>% group_by(ID, BotStrategy) %>% 
  mutate(
    'CumulativePayoff' = cumsum(Payoff)/seq_along(Payoff),
    'rate_bot' = cumsum(BotChoice)/seq_along(BotChoice),
    'prevRate_bot' = lag(rate_bot)
    )
# fix the first NA in prevRate to change (i.e. 0.5)
df$prevRate_bot = ifelse(is.na(df$prevRate_bot), 0.5, df$prevRate_bot)

```


## Explore data
```{r}

df %>% group_by(ID, BotStrategy) %>% 
  summarize(
    n()
  )
# all participants played 40 trials against both bots

df %>% group_by(BotStrategy) %>% 
  summarize(
    length(unique(ID))
  )
# All participants played both bots

```

### visualize games

for one participant only
```{r}
one_participant = df[df$ID=='Oxford',]

ggplot(one_participant, aes(Trial, CumulativePayoff, color=BotStrategy)) +
  geom_line() + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw()

```

All participants:
```{r}

ggplot(data=df, aes(x=Trial, y=CumulativePayoff, color=ID)) + 
  geom_line(size=1) + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw() + facet_wrap(.~BotStrategy)
```

```{r}
plot_df <- df %>% group_by(Trial, BotStrategy) %>% 
  summarize(
    meanPayoff = mean(CumulativePayoff, na.rm=TRUE),
    sd = sd(CumulativePayoff, na.rm=TRUE)
  )

ggplot(data=plot_df, aes(x=Trial, y=meanPayoff, color=BotStrategy)) + 
  geom_ribbon(aes(ymin=meanPayoff-sd, ymax=meanPayoff+sd), alpha=0.1, linetype=0) + 
  geom_line(size=1) + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw()

```

## Analyze data
We would expect that the participants would use a memory strategy against the random bias model and win-shift-loose-stay against the WSLS model. 
We would want to model both strategies on behavior in trials against random and WSLS bots respectively, and then compare these models to see which fits the actual behavior of the participants the best.

```{r}
# load stan models
rand_file <- file.path('multilevel_random_agent.stan')
mem_file <- file.path('multilevel_memory_agent.stan')

rand_mod <- cmdstan_model(rand_file, cpp_options = list(stan_threads=TRUE))
mem_mod <- cmdstan_model(mem_file, cpp_options = list(stan_threads=TRUE))
```



### Against random bias bot
```{r}
# ids for test
id1 = 'Oxford'
id2 = 'Hamburg'

# make data frames
df_rb <- df %>% filter(BotStrategy=="RandomBias") %>% filter(ID != id1 | ID != id2)
df_rb_test <- df %>% filter(BotStrategy=="RandomBias") %>% filter(ID == id1 | ID == id2)

# dataframe with choice and previous rate of the bot
df_h <- df_rb %>% subset(select=c(ID, Choice)) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(names_from = ID, values_from = Choice)

df_prevRate <- df_rb %>% subset(select=c(ID, prevRate_bot)) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(names_from = ID, values_from = prevRate_bot)

df_h_test <- df_rb_test %>% subset(select=c(ID, Choice)) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(names_from = ID, values_from = Choice)

df_prevRate_test <- df_rb_test %>% subset(select=c(ID, prevRate_bot)) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(names_from = ID, values_from = prevRate_bot)

```

**Memory model**
```{r}

n_trials <- max(df_rb$Trial)
n_agents <- length(unique(df_rb$ID))
n_agents_test = length(unique(df_rb_test$ID))

# data for this model
mem_data <- list(trials = n_trials, 
                 agents = n_agents,
                 h = as.matrix(df_h[, 2:(n_agents+1)]), 
                 prevRate = as.matrix(df_prevRate[, 2:(n_agents+1)]),
                 agents_test = n_agents_test,
                 h_test = as.matrix(df_h_test[, 2:(n_agents_test+1)]), 
                 prevRate_test = as.matrix(df_prevRate_test[, 2:(n_agents_test+1)])
)

# following command calls stan
mem_samples <- mem_mod$sample(
  data = mem_data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)

```

```{r}
# output of memory model
mem_samples

draws_df <- as_draws_df(mem_samples$draws())
```

```{r}
# prior and posterior plots
ggplot(draws_df) +
  geom_density(aes(alphaM), fill="blue", alpha=0.3) +
  geom_density(aes(alphaM_prior), fill="red", alpha=0.3) + 
  xlab("AlphaM") +ylab("Posterior Density") +
  theme_bw()

ggplot(draws_df) +
  geom_density(aes(alphaSD), fill="blue", alpha=0.3) +
  geom_density(aes(alphaSD_prior), fill="red", alpha=0.3) + 
  xlab("AlphaSD") +ylab("Posterior Density") +
  theme_bw()

ggplot(draws_df) +
  geom_density(aes(betaM), fill="blue", alpha=0.3) +
  geom_density(aes(betaM_prior), fill="red", alpha=0.3) + 
  xlab("BetaM") +ylab("Posterior Density") +
  theme_bw()

ggplot(draws_df) +
  geom_density(aes(betaSD), fill="blue", alpha=0.3) +
  geom_density(aes(betaSD_prior), fill="red", alpha=0.3) + 
  xlab("BetaSD") +ylab("Posterior Density") +
  theme_bw()
```

**Random model**
```{r}
n_trials <- max(df_rb$Trial)
n_agents <- length(unique(df_rb$ID))
n_agents_test = length(unique(df_rb_test$ID))

# data for this model
rand_data <- list(trials = n_trials, 
                 agents = n_agents,
                 h = as.matrix(df_h[, 2:(n_agents+1)]), 
                 agents_test = n_agents_test,
                 h_test = as.matrix(df_h_test[, 2:(n_agents_test+1)])
)

# following command calls stan
rand_samples <- rand_mod$sample(
  data = rand_data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)
```

```{r}
# output of random bias model
rand_samples

draws_df <- as_draws_df(rand_samples$draws())
```

```{r}
# prior and posterior plots
ggplot(draws_df) +
  geom_density(aes(thetaM), fill="blue", alpha=0.3) +
  geom_density(aes(thetaM_prior), fill="red", alpha=0.3) + 
  xlab("ThetaM") +ylab("Posterior Density") +
  theme_bw()

ggplot(draws_df) +
  geom_density(aes(thetaSD), fill="blue", alpha=0.3) +
  geom_density(aes(thetaSD_prior), fill="red", alpha=0.3) + 
  xlab("ThetaSD") +ylab("Posterior Density") +
  theme_bw()

```

**model comparison**
```{r}
loo_mem <-mem_samples$loo(cores=3)
loo_rand <- rand_samples$loo(cores=3)

loo_compare(loo_mem, loo_rand)

```

