---
title: "assignment n"
author: "Sara Østergaard"
date: "16/3/2022"
output: html_document
---

During this assignment you will have to pick two (or more) models (e.g. Random Bias, Imperfect Memory, WSLS) and analyze one the Matching Pennies datasets available for the course (see end of instructions)

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
pacman::p_load(
  tidyverse,
  here,
  posterior,
  cmdstanr,
  brms, 
  boot
)
```

## load data
Data with cogsci students
```{r}
df <- read_csv('mp_students_22.csv')%>% 
  filter(BotStrategyN == -1 | BotStrategyN == -2) %>%   # we will only look at WSLS and random bias bots
  mutate(BotStrategy = ifelse(BotStrategyN == -1, 'RandomBias', 'WSLS'))

df <- df %>% group_by(ID, BotStrategy) %>% 
  mutate(
    'CumulativePayoff' = cumsum(Payoff)/seq_along(Payoff),
    'rate_bot' = cumsum(BotChoice)/seq_along(BotChoice)
    )

# create col for previous rate for the bot (used in modeling the memory strategy)
prevRate_bot <- c()
for (i in 1:nrow(df)){
  if (i==1){
    prevRate_bot <- c(prevRate_bot, 0.5)
  } else {
      prevRate_bot <- c(prevRate_bot, df$rate_bot[i-1])
    }
  
}
df$prevRate_bot <- prevRate_bot
```


## Explore data
```{r}

df %>% group_by(ID, BotStrategy) %>% 
  summarize(
    n()
  )
# all participants played 40 trials against both bots

df %>% group_by(BotStrategy) %>% 
  summarize(
    length(unique(ID))
  )
# All participants played both bots

```

### visualize games

for one participant only
```{r}
one_participant = df[df$ID=='Oxford',]

ggplot(one_participant, aes(Trial, CumulativePayoff, color=BotStrategy)) +
  geom_line() + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw()

```

All participants:
```{r}

ggplot(data=df, aes(x=Trial, y=CumulativePayoff, color=ID)) + 
  geom_line(size=1) + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw() + facet_wrap(.~BotStrategy)
```

```{r}
plot_df <- df %>% group_by(Trial, BotStrategy) %>% 
  summarize(
    meanPayoff = mean(CumulativePayoff, na.rm=TRUE),
    sd = sd(CumulativePayoff, na.rm=TRUE)
  )

ggplot(data=plot_df, aes(x=Trial, y=meanPayoff, color=BotStrategy)) + 
  geom_ribbon(aes(ymin=meanPayoff-sd, ymax=meanPayoff+sd), alpha=0.1, linetype=0) + 
  geom_line(size=1) + 
  geom_hline(yintercept=0, linetype='dashed') +
  theme_bw()

```

## Analyze data
We would expect that the participants would use a memory strategy against the random bias model and win-shift-loose-stay against the WSLS model. 
We will model both strategies on behavior in trials against random and WSLS bots respectively, and then compare these models to see which fits the actual behavior of the participants the best.

```{r}
# load stan models
mem_file <- file.path('../a2/memory_agent.stan')
wsls_file <- file.path('../a2/WSLS_agent.stan')

mem_mod <- cmdstan_model(mem_file, cpp_options = list(stan_threads=TRUE))
wsls_mod <- cmdstan_model(wsls_file, cpp_options = list(stan_threads=TRUE))
```

### Against random bias bot
```{r}
# make data frame
df_rb <- df %>% filter(BotStrategy=="RandomBias")

```

```{r}
# data for this model
mem_data <- list(n = nrow(df_rb), h = df_rb$Choice, r = df_rb$prevRate_bot,
             prior_mean_alpha=0, prior_mean_beta=0, prior_sd_alpha=1, prior_sd_beta=0.5)

# following command calls stan
samples <- mem_mod$sample(
  data = mem_data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)

```

```{r}
# data for this model
wsls_data <- list()

# following command calls stan
samples <- wsls_mod$sample(
  data = wsls_data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)
```

