---
title: "Assignment 1"
author: "Studygroup: Johan, Simon, Emil & Sara"
date: "16/2/2022"
output: html_document
---

Describe and motivate two possible verbal model of how the Matching Pennies Task is solved and provide an algorithmic formalization (in R). Set them to play against randomAgent and WSLSAgent (120 trials, 100 agents). Visualize and comment the simulation.

The algorithmic formalization fo the agent strategies are in agent_strategies.R, and the functions for having the agents play against each other are in games.R.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
pacman::p_load(
  tidyverse
)

source('agent_strategies.R')
source('games.R')

```

## Random Agent
```{r}
trials <- 125
d2 <- NULL

for (noise in seq(0,0.5,0.1)){
  for (rate in seq(0,1,0.1)){
    randomChoice <- rep(NA, trials)
    
    for (t in seq(trials)){
      randomChoice[t] <- RandomAgent(rate, noise)
    }
    
    temp <- tibble(
      trial = seq(trials),
      choice = randomChoice,
      rate,
      noise
    )
    if (is.null(d2)){
      d2 <- temp
    } else {
      d2 <- rbind(d2,temp)
    }
  }
}
```

```{r}
d2 <- d2 %>% group_by(rate, noise) %>% 
  mutate(
    cumulative = cumsum(choice) /seq_along(choice)
  )

ggplot(d2, aes(trial, cumulative, group=rate, color=rate)) +
  geom_line() + 
  geom_hline(yintercept=0.5, linetype='dashed') +
  ylim(0,1) + 
  facet_wrap(.~noise) + theme_bw()
```

## RandomAgent (mismatcher) against memoryAgent (matcher)

One round
```{r}
trials <- 125
rate <- 0.5
noise <- 0.2
mem_constraint <- 5

df <- RandomAgainstMemory(trials, rate, noise, mem_constraint)

```


```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, matcher)) +  # matcher
  geom_line(color='blue', aes(trial, mismatcher)) +  # mismatcher
  theme_bw()
```

```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, cumulativeMatcher)) +  # matcher
  geom_line(color='blue', aes(trial, cumulativeMismatcher)) +  # mismatcher
  theme_bw()
```
100 rounds with different parameter values
```{r}
# j <- 0
# for (i in 1:100){  # number of repetions
#   for(r in seq(0,1,0.2)){  # rate
#     for (n in seq(0,0.5,0.2)){  # noise
#       tmp <- RandomAgainstMemory(
#         trials = 120,
#         rate = r,
#         noise = n,
#         mem_constraint= 120
#       )
# 
#       tmp$rate <- r
#       tmp$noise <- n
#       tmp$mem_constraint <- m
#       tmp$rep <- i
# 
#       if (i ==1){ # first round
#         df <- tmp
#       } else {
#         df <- rbind(df, tmp)
#       }
# 
#       # print statement
#       if (i != j){
#         print(paste('Round:', i))
#       }
#       j <- i
#     }
#   }
# }
# 
# write_csv(df, "RandomAgainstMemory.csv")
```

```{r}
df <- read_csv('RandomAgainstMemory.csv')

df %>% group_by(rate, mem_constraint, noise) %>% 
  summarize(
    mean(Feedback)
  )
```
```{r}
df_plot <- df %>% group_by(trial, noise, rate) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback))
  )

ggplot(data=df_plot, aes(x=trial, y=cumulativeMatcher, color=as.factor(noise))) + 
  geom_ribbon(aes(ymin=cumulativeMatcher-sd, ymax=cumulativeMatcher+sd), alpha=0.1, linetype=0) + 
  geom_line() +  
  facet_wrap(.~rate) + 
  theme_bw() + labs(title = 'Memory agent (matcher) against Random agent by rate', color='Noise') + scale_colour_brewer(palette = 'Accent')
```

## WSLSAgent (mismatcher) against MemoryAgent (matcher)

One round
```{r}
trials <- 125
noise <- 0.0
mem_constraint <- 125

df <- WSLSAgainstMemory(trials, noise, mem_constraint)
```

```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, matcher)) +  # matcher
  geom_line(color='blue', aes(trial, mismatcher)) +  # mismatcher
  theme_bw()
```

```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, cumulativeMatcher)) +  # matcher
  geom_line(color='blue', aes(trial, cumulativeMismatcher)) +  # mismatcher
  theme_bw()
```

100 rounds with different parameter values
```{r}
# j <- 0
# for (i in 1:100){  # number of repetions
#   for (n in seq(0,0.5,0.2)){  # noise
#     for (m in seq(5,125,30)){  # memory constraint
#       tmp <- WSLSAgainstMemory(
#         trials = 120,
#         noise = n,
#         mem_constraint= m
#       )
# 
#       tmp$noise <- n
#       tmp$mem_constraint <- m
#       tmp$rep <- i
# 
#       if (i ==1){ # first round
#         df <- tmp
#       } else {
#         df <- rbind(df, tmp)
#       }
# 
#       # print statement
#       if (i != j){
#         print(paste('Round:', i))
#       }
#       j <- i
#     }
#   }
# }
# 
# write_csv(df, "WSLSAgainstMemory.csv")
```

```{r}
df <- read_csv('WSLSAgainstMemory.csv')

df %>% group_by(mem_constraint, noise) %>% 
  summarize(
    mean(Feedback),
    sd(Feedback)
  )
```

```{r}
df_plot <- df %>% group_by(trial, noise) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback))
  )

ggplot(data=df_plot, aes(x=trial, y=cumulativeMatcher, color=as.factor(noise))) + 
  geom_ribbon(aes(ymin=cumulativeMatcher-sd, ymax=cumulativeMatcher+sd), alpha=0.1, linetype=0) + 
  geom_line(size=1) + 
  theme_bw() + labs(title = 'Memory agent (matcher) against WSLS probability agent', color='Noise') + scale_color_brewer(palette = 'Accent')
```


## WSLS prob (mismatcher) against memory (matcher)

One round
```{r}
trials <- 125
noise <- 0.0
mem_constraint <- 125
prob <- 0.8

df <- WSLSpAgainstMemory(trials, noise, mem_constraint, prob)

```

```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, matcher)) +  # matcher
  geom_line(color='blue', aes(trial, mismatcher)) +  # mismatcher
  theme_bw()
```

```{r}
ggplot(df) + 
  geom_line(color='red', aes(trial, cumulativeMatcher)) +  # matcher
  geom_line(color='blue', aes(trial, cumulativeMismatcher)) +  # mismatcher
  theme_bw()
```

100 rounds with different parameter values
```{r}
# j <- 0
# for (i in 1:100){  # number of repetions
#   for (n in seq(0,0.5,0.2)){  # noise
#     for (m in seq(5,125,30)){  # memory constraint
#       tmp <- WSLSpAgainstMemory(
#         trials = 120,
#         noise = n,
#         mem_constraint= m,
#         prob=0.8
#       )
# 
#       tmp$noise <- n
#       tmp$mem_constraint <- m
#       tmp$rep <- i
# 
#       if (i ==1){ # first round
#         df <- tmp
#       } else {
#         df <- rbind(df, tmp)
#       }
# 
#       # print statement
#       if (i != j){
#         print(paste('Round:', i))
#       }
#       j <- i
#     }
#   }
# }
# 
# write_csv(df, "WSLSprobAgainstMemory.csv")
```

```{r}
df <- read_csv('WSLSprobAgainstMemory.csv')

df %>% group_by(mem_constraint, noise) %>% 
  summarize(
    mean(Feedback),
    sd(Feedback)
  )
```

```{r}
df_plot <- df %>% group_by(trial, noise) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback)),
  )

ggplot(data=df_plot, aes(x=trial, y=cumulativeMatcher, color=as.factor(noise))) + 
  geom_ribbon(aes(ymin=cumulativeMatcher-sd, ymax=cumulativeMatcher+sd), alpha=0.1, linetype=0) + 
  geom_line(size=1) + 
  theme_bw() + labs(title = 'Memory agent (matcher) against WSLS probability agent', color='Noise') + scale_color_brewer(palette = 'Accent')
```



## Compare different strategies against memory
```{r}
df_random <- read_csv('RandomAgainstMemory.csv') %>% group_by(trial) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback)),
    opponent = 'Random',
    player = 'Memory'
  )

df_WSLS <- read_csv('WSLSAgainstMemory.csv') %>% group_by(trial) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback)),
    opponent = 'WSLS',
    player = 'Memory'
  )

df_WSLSp <- read_csv('WSLSprobAgainstMemory.csv') %>% group_by(trial) %>% 
  summarize(
    cumulativeMatcher = mean(cumsum(Feedback)/seq_along(Feedback)),
    sd = sd(cumsum(Feedback)/seq_along(Feedback)),
    opponent = 'WSLS_prob',
    player = 'Memory'
  )

df_plot <- rbind(df_random, df_WSLS, df_WSLSp)
```

```{r}
ggplot(data=df_plot, aes(x=trial, y=cumulativeMatcher, color=opponent)) + 
  geom_ribbon(aes(ymin=cumulativeMatcher-sd, ymax=cumulativeMatcher+sd), alpha=0.1, linetype=0) + 
  geom_line(size=1) + 
  theme_bw() + labs(title = 'Memory (matcher) agent against opponent (mismatcher)', color='Opponent') + scale_color_brewer(palette = 'Accent')
```

